Characterizing Stellar Streams by ML

### 2023-08-14 src/tex/ms.tex
[showyourwork] overleaf sync
[showyourwork] DAG
Signed-off-by: nstarman <nstarman@users.noreply.github.com>

\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{todonotes}

@@

\newcommand{\package}[1]{\textsc{#1}}
\newcommand{\stream}[1]{#1}
\newcommand{\dataarchive}[1]{\texttt{#1}}

@@

\newcommand{\cmp}[2]{#2^{(#1)}}
\newcommand{\smallcomponent}[2]{#2^{\scriptscriptstyle (#1)}}
\newcommand{\astroM}[1]{\smallcomponent{w}{#1}}
\newcommand{\photoM}[1]{\smallcomponent{m}{#1}}
\newcommand{\chemoM}[1]{\smallcomponent{z}{#1}}

@@

\newcommand{\JN}[1]{{\textcolor{red}{JN: #1}}}

@@

\section{Introduction} \label{sec:intro}

    \todo[inline]{Rephrase ---}

@@

        % TODO: significantly expand on this

        \paragraph{Modeling approach}

@@

        The hypertube idea is connected to orbits.
        One might think then that a variable like the phase-angle or the arc-length, or any variable that directly sweeps along the stream would be best to use as the conditioning variable.
        This is not the case for a number of reasons (see \cite{StarkmanEtAl2023} for some discussion on this topic), not least of which is that we do not have any of these quantities. We do not know the stream track, it's what we want to know.

@@

        \begin{figure}
        \centering
        \includegraphics[scale=0.5]{static/DAG_StreamDensity.pdf}
        \caption{Fill in.}
        \label{fig: DAG}
        \end{figure}

        We model the likelihood for a single star as
        \begin{align} \label{eq:general_model}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
                &= \sum_{q} \ f_q(w_{i, 0}) \pdf(\nth{\mbs{w}}|\mbs{\theta},q_i\!=\!q) \pdf(\nth{\mbs{m}}|\mbs{\theta},q_i\!=\!q)
        \end{align}
        where we have assumed that $\nth{\mbs{w}}$ and $\nth{\mbs{m}}$ are
        conditionally independent, and that the mixture weights $f$ are
        normalized s.t. $\sum_q f(w_{i,0}) = 1$.  We discuss the mixture weights
        further in \autoref{sub:mixture_weight}.

        Assuming the measurement of each star is independent, the total
        log-likelihood is a sum over all stars:
        \begin{equation}
            \ln\mcal{L}\left(\{(\nth{\mbs{w}},\nth{\mbs{m}})\} | \mbs{\theta}\right) = \sum_i  \ln \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta}).
        \end{equation}

        It will prove convenient to distinguish the models for $q=0,1$ as the
        \textit{background} and \textit{stream} models, respectively. In this
        notation:

        \begin{align}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} P^{(S)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(S)}(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) P^{(B)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(B)}(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}

        The probablity of a being a stream member is then simply
        \begin{equation}\label{eq:membership_prob}
            \pdf\left(S | \mbs{w}_n, \mbs{m}_n, \mbs{\theta} \right) = \frac{f P^{(S)}(\nth{\mbs{w}}|\mbs{\theta}) P^{(S)}(\nth{\mbs{m}}|\mbs{\theta}) }{ \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})}.
        \end{equation}

        Hereafter, when we refer to ``membership probability,"
        \autoref{eq:membership_prob} provides the formal definition.

        \paragraph{missing phase-space observations}

            Streams have a diversity of measurement coverage: several have full 6-D astrometrics as well as photometrics and chemical abundances available \citep[e.g.,][]{Koposov2019, 2020A&A...635L...3A, 2022ApJ...928...30L}, while most have partial coverage, lacking feature dimensions.
            % TODO: terminology ND-phase-space vs feature columns
            Most streams in \dataarchive{Gaia} {\color{red} [cite]}, for instance, have proper motions but not radial velocities nor reliable parallax measurements. Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many feature dimensions, however only for a select set of streams. Given the diversity of coverage it is important that the likelihood model can deal with missing data.

            The method of treatment of missing data depends on the cause for that data to be missing.
            Broadly, there are two categories of causes: randomness and systematics.
            Consider a dataset with all features measured but with a random subset of features masked.
            One possible approach is to attempt to impute the missing features, then feed the full-featured data to the likelihood model. Alternatively, the likelihood model might use only the present features, ignoring those missing. If however, features are missing for systematic reasons, then this approach can introduce systematic bias. Consider a star too far away to measure it's parallax distance or radial velocity. This is a systematic reason and this information should be incorporated into any model as a prior. Not doing so, the data imputation approach might predict too-close distances. Likewise, using only present features in the likelihood model misses that the star is likely very distant.

            {\color{red} Sentence explaining why we ignore features regardless.}
            We include a quality flag on our calculations, indicating the number of missing features.

            The per-star likelihood is modelled with a distribution with the same dimensionality as the data vector.
            For example, a star with a 6D astrometric measurement will be modelled with a full 6D distribution, while a star missing its radial velocity measurement, will be modelled with a 5D dimensionally-reduced form of the 6D distribution. Using dimensionally reduced distributions ensure that missing phase-space dimensions do not amplify nor suppress the likelihood.

@@

        In particular, for a given field the distribution of $\phi_1$ coordinates is specified by the set $\{\nth{\phi_{1}}\}_f$. A normalizng flow is trained to this dataset, and will satisfy $\int d\phi_1 P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of stars in the field considered. We can then obtain the linear stream density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q \left[N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) \right] .
        \end{equation}

@@

            The astrometric track is modeled with a neural network, encompassing each phase space dimension, parameterized as a function of $\phi_1$. The neural network output for a single stream component is the mean location of the stream track in each dimension, $\astroM{\mbs{\mu}}(\phi_1)$, and its associated covariance matrix $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is then
            \begin{equation}
                \pdf^{(S)}(\nth{\mbs{w}} | \astroM{\mbs{\theta}}(\nth{\phi_{1}})) = \mcal{N}_F \big(\nth{\mbs{w}} | \astroM{\mbs{\mu}} \left(\nth{\phi_{1}} \right), \astroM{\mbs{\Sigma}} \left(\nth{\phi_{1}}\right) \big),
            \end{equation}
            where $\mbs{\theta}$ is the set of neural network parameters that define $\astroM{\mbs{\mu}}$ and $\astroM{\mbs{\Sigma}}$.
            In the simplest case, we take $\astroM{\mbs{\Sigma}}$ to be diagonal, consisting of both intrinsic dispersion for each dimension of $\mbs{w}$ and the data uncertainties added in quadrature.

@@

            In the presence of missing phase-space coordinate $f_k$ we reduce the dimensionality of the Gaussian to that of the data-vector. Namely, for $\mbs{I}_{\tilde{F}} = \mbs{I}_{F} \backslash f_k$
            \begin{align}
                \nth{\tilde{\mbf{w}}} &\mapsto \nth{\tilde{\mbf{w}}} = \{ w_{n,f}\}_{f\in \mbs{I}_{\tilde{F}}}, \nonumber\\
                \astroM{\mbs{\mu}} &\mapsto \astroM{\tilde{\mbs{\mu}}} = \{ \mu_{f}\}_{f\in\mbs{I}_{\tilde{F}}}, \nonumber\\
                \astroM{\mbs{\Sigma}} &\mapsto \astroM{\tilde{\mbs{\Sigma}}} = \{ \Sigma_{i,j}\}_{i,j \in \mbs{I}_{\tilde{F}}}, & \nonumber
                \intertext{such that,}
                \pdf^{(S)}(\nth{\mbs{w}} & | \astroM{\mbs{\theta}}(\nth{\phi_{1}}))
                    \mapsto \pdf^{(S)}(\nth{\tilde{\mbs{w}}} | \astroM{\tilde{\mbs{\theta}}} (\nth{\phi_{1}})).
            \end{align}

@@

            Let $\fth{\cmp{B}{\mbs{\theta}}}(\phi_1)$ represent the neural network which outputs a vector of parameters characterizing the background distribution at a given $\phi_1$ for astrometric dimension $f$. The number of outputs for the background model is equal to $\mathrm{dim}\left(\fth{\cmp{B}{\mbs{\theta}}}\right)$ (where $f=0$ corresponds to $\phi_1$, which we do not model). At a given $\phi_1$, the probability density for the astrometric dimension $f>0$ is $\fth{\pdf^{(B)}}\left(\mbs{w}_{n,f} | \fth{\cmp{B}{\mbs{\theta}}}(\phi_1) \right)$. We treat each background component as independent, allowing us to write the likelihood as a product over the astrometric dimensions:
            \begin{equation} \label{eq:astrometric_model_off_stream_probability}
                \pdf^{(B)}\left(\nth{\mbs{w}} | \mbs{\theta}(\phi_1) \right) = \prod_{f\in \mbs{I}_{\tilde{F}}} \pdf_f^{(B)}\left(\mbs{w}_{n,f} | \fth{\cmp{B}{\mbs{\theta}}}(\phi_1) \right).
            \end{equation}
            Where $\mbs{I}_{\tilde{F}}$ is the set of indices of the features, omitting missing phase space dimensions.

    %\newpage

@@

            At any given $\phi_1$, the isochrone is modeled in absolute magnitudes
            and parameterized by the unit-normalized stellar mass, labeled $\gamma$. For example, increasing $\gamma$ corresponds to movement up the main sequence towards the red giant branch. We denote the isochrone as $\mcal{I(\gamma)}$. The intrinsic dipsersion around the isochrone is $\Sigma_\mcal{I}(\gamma)$. Since the isochrone is in absolute magnitudes, but the data are in apparent magnitudes,
            it is necessary to shift the isochrone by a distance modulus to the predicted distance of the stream, labeled  $\mu(\nth{\phi_1})$.

            For the $n$-th star, the photometric model is:

            \begin{multline}
                P^{(S)}(\mbs{m_i}, \gamma; \mbs{\theta}(\nth{\phi_1}), \nth{\phi_1}) 
                \\ \sim \mcal{N}(\nth{\mbs{m}}; \{\mcal{I(\gamma)} + \mu(\nth{\phi_1}), \sqrt{\Sigma_{\mcal{I}}^2(\gamma) + \Sigma_\mu^2(\nth{\phi_1}) + \Sigma_i^2} \}) \\ \times\pdf(\gamma, \nth{\phi_1}),
            \end{multline}

            where $\pdf(\gamma)$ encodes both the present-day mass function of the stream, as well as the observational constraints (e.g. $g < 22$ mag).

            We are interested in the total probability (at some $\phi_1$) and so integrate over $\gamma$ to find:
            \begin{small}
            \begin{multline}
                P^{(S)}(\nth{\mbs{m}}; \mbs{\theta}(\nth{\phi_1}), \nth{\phi_1}) = \\\int\limits_{\gamma=0}^{1} \mcal{N}(\mbs{m}; \{\mcal{I(\gamma)} + \mu(\nth{\phi_1}), \sqrt{\Sigma_{\mcal{I}}^2(\gamma) + \Sigma_\mu^2(\nth{\phi_1}) + \Sigma_i^2} \}) \pdf(\gamma) \rm{d}\gamma.
            \end{multline}\end{small}
            Then, for a given $\phi_1$ we can obtain a distribution over the distance modulus using 
            \begin{equation}
                \mu_{\rm sample}\left(\phi_1\right) \sim \mathcal{N}\left(\mu(\phi_1), \Sigma^2_\mu(\phi_1) \right) \quad\unit{mag}.
            \end{equation}
            For each sample, we can convert the distance-modulus distribution to the distance track using 
            \begin{equation}
                \rm{dist}\left(\phi_1\right) = 10^{\frac{1}{5}\mu_{\rm sample}(\phi_1)+1} \quad\unit{pc}.
            \end{equation}

            \todo[inline]{TODO:}
            \begin{enumerate}
                \item Not including isochrone parameters, e.g. age and metallicity
                \item integrating over $\gamma$ the model assumes a PDMF, e.g. Kroupa
                \item generic details of neural network, e.g Table 1
                \item introducing parallax
                \item how we reduce dimensionality.
            \end{enumerate}

>>>>

\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{dsfont}
\usepackage{todonotes}

@@

\newcommand{\package}[1]{\textsc{#1}}
\newcommand{\stream}[1]{#1}
\newcommand{\dataarchive}[1]{\texttt{#1}}
\newcommand{\Gaia}{\dataarchive{Gaia}}
\newcommand{\gaia}{\dataarchive{gaia}}

@@

\newcommand{\smallcomponent}[2]{#2^{\scriptscriptstyle (#1)}}
\newcommand{\cmp}[2]{\smallcomponent{#1}{#2}}
\newcommand{\Scmp}[1]{\cmp{S}{#1}}
\newcommand{\Bcmp}[1]{\cmp{B}{#1}}
\newcommand{\astroM}[1]{{#1}_w}
\newcommand{\photoM}[1]{\smallcomponent{m}{#1}}
\newcommand{\chemoM}[1]{\smallcomponent{z}{#1}}

\newcommand{\Spdf}{\Scmp{\pdf}}
\newcommand{\Bpdf}{\Bcmp{\pdf}}

@@

\newcommand{\TODO}[1]{{\textcolor{red}{#1}}}
\newcommand{\JN}[1]{\TODO{#1}}

@@

\section{Introduction} \label{sec:intro}
    Stellar streams are the disrupted remnants of globular clusters and satellite galaxies, and their census has grown extensively in the Milky Way thanks to all-sky astrometric missions like {\it Gaia} and low-surface brightness photometric surveys (\JN{cite}). The formation of streams can be traced back to a progenitor cluster (i.e., a satellite galaxy or globular cluster), whose stars are gradually lost to the tidal field of the host galaxy and extend along a series of similar orbits producing tidal tails which can sometimes span many tens of degrees across the sky. This orbital proximity of stars belonging to a stream makes them sensitive detectors of the mass distribution of galaxies, sourced by both the baryonic and dark matter components across a large range of mass-scales.

    The average properties of a stream (e.g., the mean position and velocity track of the tidal tails) can be used to reconstruct the global dark matter distribution of the stellar halo (e.g., \JN{cite}), while local variations in the stream track and surface density encode information about the stream's encounters with baryonic structures such as (e.g.) the stellar bar (\JN{cite}), and non-luminous structures like dark matter subhalos (\JN{cite}).

    While several methods have been developed to model stream properties as a function of global halo properties (e.g., flattening and orientation;\JN{cite}) and local substructures (e.g., the bar's pattern speed and the mass of subhalo perturbers; \JN{cite}), these methods and insights often rely on the existence of a homogenized stream-tracer population that characterizes the stream across each of the observable phase-space dimensions. Additionally, because even the fine-grain properties of streams are influenced by the detailed mass distribution of galaxies, it is important to propagate errors both in measurement uncertainty and membership uncertainty (i.e., which stars belong to the stream versus the background) when attempting to model the galaxy. Otherwise, a stream that has been poorly characterized could bias constraints on (e.g.) the total mass of the galaxy, the dark matter halo shape, and the mass of subhalo perturbers. In order to generate more precise constraints on the structure of the galaxy, it is therefore imperative that a statistically sound and homogenized catalog of stellar streams is produced so that theoretical models for the gravitational potential of the galaxy across a large range of spatial scales can be constrained.

@@

@@

        {\color{red} Assume we know a stream track}

@@

        \IfFileExists{static/DAG_StreamDensity.pdf}{
            \begin{figure}
                \centering
                \includegraphics[scale=0.5]{static/DAG_StreamDensity.pdf}
                \caption{Fill in.}
                \label{fig:DAG}
            \end{figure}
        }{
        }

        Mixture models (MMs) are a statistical method to represent a population as a set of a sub-populations specified by probability distributions. These distributions are characterized by their parameters and the mixture by a corresponding set of mixture coefficients. Scalar mixture models may be extended to model completely general conditional distributions by making the distribution parameters and mixture coefficients functions of a conditional input \citep{McLachlan1989MixtureM}.
        Mixture Density Networks (MDNs) are a machine learning method to implement these general MM, taking the parameters and coefficients to be the outputs of feed-forward neural networks and as such, general continuous functions \citep{Bishop1994}. Functional parameters allows the model to capture variation over those parameters that is not possible with scalar-valued models. Moreover, by using neural networks the inferred functional forms are ``model-free" and driven entirely by the data.

        For the MDN we take $x_{\cdot, 0}$, which is the first feature column of the data $\mbf{X}$, as the `independent' coordinate over which the MDN parameters are conditioned. In the context of streams $x_{\cdot, 0} \equiv \phi_1$ is the longitude coordinate in a reference frame rotated such that the stream is aligned along $\phi_1$. The parameters of the stream, the background field, and the mixing coefficients of the two are now general data-driven continuous functions along $\phi_1$.
        {\color{cyan} \\ connecting to intro...

            The hypertube idea is connected to orbits.
            One might think then that a variable like the phase-angle or the arc-length, or any variable that directly sweeps along the stream would be best to use as the conditioning variable.
            This is not the case for a number of reasons (see \cite{StarkmanEtAl2023} for some discussion on this topic), not least of which is that we do not have any of these quantities. We do not know the stream track, it's what we want to know.
        }

        The PDF of a general MDN is
        \begin{equation} \label{eq:general_mixture_network}
            \!\!\! \pdf(\nth{\mbs{x}} | \mbs{\theta}(\phi_1))
            \!=\! \sum_{\mrm{q} \in \mbs{I_Q}} \! \qth{f}(\phi_1) \cmp{q}{\pdf}(\nth{\mbs{x}}|\qth{\mbs{\theta}}(\phi_1))
            % \prior(\nth{\mbs{x}}, \phi_1)
        \end{equation}
        where $\pdf_q$ is the PDF of the $q$-th model and $f_q$ its mixing coefficient for all $q$ models in the set of models indexed by $\mbs{I_Q}$. $\mbs{\theta}$ is the $\phi_1$-conditional output of the feed-forward neural network, and $\nth{\mbs{x}}$ the data on the $n$-th star in $\mbs{X}$, not including $\phi_1$. For convenience we drop the implied subscript $n$ in $\phi_1$. A full detailing of notation and symbols is included in \autoref{app:notation}. For convenience we drop the explicit notation of $\mbs{\theta}$'s dependence on $\phi_1$ and the index set $\mbs{I}_X$ in summations over its elements $x$.

        The mixture coefficients $\qth{f}$ are normalized s.t.
        \begin{equation}
            \sum_{\mrm{q}} \qth{f}(\phi_1) = 1 \qquad \forall \phi_1
        \end{equation}
        In practice this is enforced by defining
        \begin{equation} \label{eq:mixture_weight_normalization}
            f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash \{b\}} \qth{f}(\phi_1),
        \end{equation}
        the summation over all models except for some ``background" index $b \in \mbs{I_Q}$.
        We discuss the mixture weights further in \autoref{sub:mixture_weight}.

        In this work we consider primarily astrometric and photometric data, (notated $\mbs{W}$ and $\mbs{M}$, respectively, as detailed in \autoref{app:notation}), but note that the MDN framework may be extended to arbitrary feature dimensions.
        Assuming the conditional independence of the astrometry and photometry,
        we split each model into linearly separable astrometric and photometric models $\cmp{q}{\pdf} = \cmp{q,w}{\pdf} \cmp{q,m}{\pdf}$ .

        The likelihood for a single star then becomes, simplifying from \autoref{eq:general_mixture_network},
        \begin{equation} \label{eq:general_model}
            \!\!\!\!\! \pdf(\nth{\mbs{w}}, \! \nth{\mbs{m}} | \mbs{\theta})
                \!=\! \sum_{q} \! f_q(\phi_1) \cmp{q,w}{\pdf}(\nth{\mbs{w}}|\mbs{\theta}) \cmp{q,m}{\pdf}(\nth{\mbs{m}}|\mbs{\theta})
        \end{equation}

        Assuming the measurement of each star is independent, the total
        log-likelihood is a sum over all stars:
        \begin{equation} \label{eq:general_likelihood}
            \ln\mcal{L}\left(\{\nth{\mbs{w}},\nth{\mbs{m}}\} | \mbs{\theta}\right) = \sum_n \ln \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta}).
        \end{equation}

        It will prove convenient to distinguish the models for $q$ as the
        \textit{background} and \textit{stream} models, respectively. In this
        notation:
        \begin{align} \label{eq:stream_and_bkg_prob}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} \Scmp{P}(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) \Bpdf(\nth{\mbs{w}}|\mbs{\theta}) \Bpdf(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}
        where we have dropped the $(w), (m)$ superscript as the model types are evident from their inputs. We applied \autoref{eq:mixture_weight_normalization} to require only a single mixture coefficient.
        Of course, if more components are included in the mixture then \autoref{eq:stream_and_bkg_prob} becomes the more general \autoref{eq:general_model}. However, this an implementation detail as any more complex model, e.g a model including a cold stream, extra-tidal features, the background stellar halo, and field globular clusters may be rearranged such that all stream-related components (cold or extra-tidal) and all other components (stellar halo and globular clusters) are grouped into multi-component \textit{stream} and \textit{background} models which themselves adhere to a \autoref{eq:general_model}-type relation.

        The probability of a being a stream member is then simply
        \begin{equation}\label{eq:membership_prob}
            \pdf\left(S | \nth{\mbs{w}}, \nth{\mbs{m}}, \mbs{\theta} \right) = \frac{f \Spdf(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) }{ \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})}.
        \end{equation}

        Hereafter, when we refer to ``membership probability,"
        \autoref{eq:membership_prob} provides the formal definition. We nuance this definition by considering missing feature dimensions in the following subsection.

    \subsection{Missing Phase-Space Observations} \label{sub:missing_data}

        Streams have a diversity of measurement coverage: several have full 6-D astrometrics as well as photometrics and chemical abundances available \citep[e.g.,][]{Koposov2019, 2020A&A...635L...3A, 2022ApJ...928...30L}, while most have partial coverage, lacking some feature dimensions.
        % TODO: terminology ND-phase-space vs feature columns
        Most streams in \gaia {\color{red} [cite]}, for instance, have proper motions but not radial velocities nor reliable parallax measurements. Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many feature dimensions, however only for a select set of streams. Given the diversity of coverage it is important that the likelihood model -- \autoref{eq:membership_prob} -- account for missing data.

        The method of treatment of missing data depends on the cause for that data to be missing.
        Broadly, there are two categories of causes: randomness and systematics.
        Consider a dataset with all features measured but with a random subset of features masked.
        One possible approach is to attempt to impute the missing features, then feed the full-featured data to the likelihood model. Alternatively, the likelihood model might use only the present features, ignoring those missing. If however, features are missing for systematic reasons, then this approach can introduce systematic bias. Consider a star too far away to measure it's parallax distance or radial velocity. This is a systematic reason and this information should be incorporated into any model as a prior. Not doing so, the data imputation approach might predict too-close distances. Likewise, using only present features in the likelihood model misses that the star is likely very distant.
        The effects of mask systematics are mitigated by not including such data,
        moving the contribution of the systematics from a prior to the evidence.
        Our models compute likelihoods under a specific dataset, and \autoref{eq:membership_prob} is a ratio of the probabilities, so the removed systematics are not expected to be impactful.

        We adopt the approach of modeling the data masks as randomly-distributed and not systematic. We do not impute missing data, instead the per-star likelihood is modelled with a distribution with the same dimensionality as the data vector.
        For example, a star with a 6D astrometric measurement will be modelled with a full 6D distribution, while a star missing its radial velocity measurement will be modelled with a dimensionally-reduced (5D) form of the 6D distribution. Using dimensionally reduced distributions ensure that missing phase-space dimensions do not amplify nor suppress the likelihood. 
        Notwithstanding, we include a quality flag on our calculations indicating the number of missing features.
        We leave incorporating priors on the data mask distribution to future work.

@@

        In particular, for a given field the distribution of $\phi_1$ coordinates is specified by the set $\{\phi_1\}_f$. A normalizng flow is trained to this dataset, and will satisfy $\int d\phi_1 P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of stars in the field considered. We can then obtain the linear stream density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q \left[N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) \right] .
        \end{equation}

@@

            The astrometric track is modeled with a neural network, encompassing each phase space dimension, parameterized as a function of $\phi_1$. The neural network output for a single stream component is the mean location of the stream track in each dimension, $\astroM{\mbs{\mu}}(\phi_1)$, and its associated covariance matrix $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is then
            \begin{equation}
                \cmp{S,w}{\pdf}(\nth{\mbs{w}} | \mbs{\theta}) \! = \mcal{N}_F \big(\nth{\mbs{w}} | \astroM{\mbs{\mu}}\!(\phi_1), {\cmp{w}{\nth{\mbs{\Sigma}}}}^2 \! + \!  \astroM{\mbs{\Sigma}}^2\!\left(\phi_1\right) \!\big),
            \end{equation}
            where $\astroM{\mbs{\theta}}$ is the set of neural network parameters that define $\astroM{\mbs{\mu}}$ and $\astroM{\mbs{\Sigma}}$; and $\cmp{w}{\nth{\mbs{\Sigma}}}$ the data uncertainties.
            In the simplest case, we take $\astroM{\mbs{\Sigma}}$ to be diagonal, consisting of both intrinsic dispersion for each dimension of $\mbs{w}$ and the data uncertainties added in quadrature. The data uncertainties $\cmp{w}{\nth{\mbs{\Sigma}}}$ are drawn from $\mbs{\delta}\mbf{W}$, the dataset of astrometric uncertainties (see \autoref{app:notation}).

@@

            In the presence of missing phase-space coordinate $f_k$ we reduce the dimensionality of the Gaussian to that of the data-vector. Namely, for $\mbs{I}_{\tilde{F}} = \mbs{I}_{F} \backslash \{f_k\}$
            \begin{align}
                \nth{\tilde{\mbf{w}}} &\mapsto \nth{\tilde{\mbf{w}}} = \{ w_{n,f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\mu}} &\mapsto \astroM{\tilde{\mbs{\mu}}} = \{ \mu_{f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\Sigma}} &\mapsto \astroM{\tilde{\mbs{\Sigma}}} = \{ \Sigma_{i,j} | i,j \in \mbs{I}_{\tilde{F}} \}, & \nonumber
                \intertext{such that,}
                \Spdf(\nth{\mbs{w}} & | \astroM{\mbs{\theta}}(\phi_1))
                    \mapsto \Spdf(\nth{\tilde{\mbs{w}}} | \astroM{\tilde{\mbs{\theta}}} (\phi_1)) = \mcal{N}_{\tilde{F}}.
            \end{align}

        % ssub: on_stream (end)

@@

            Let $\Bcmp{\fth{\mbs{\theta}}}(\phi_1)$ represent the neural network which outputs a vector of parameters characterizing the background distribution at a given $\phi_1$ for astrometric dimension $f$. The number of outputs for the background model is equal to $\mathrm{dim}\left(\fth{\Bcmp{\mbs{\theta}}}\right)$ (where $f=0$ corresponds to $\phi_1$, which we do not model). At a given $\phi_1$, the probability density for the astrometric dimension $f>0$ is $\Bcmp{\fth{\pdf}}\left(\mbs{w}_{n,f} | \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right)$. We treat each background component as independent, allowing us to write the likelihood as a product over the astrometric dimensions:
            \begin{equation} \label{eq:astrometric_model_off_stream_probability}
                \pdf^{(B)}\left(\nth{\mbs{w}} | \mbs{\theta}(\phi_1) \right) = \prod_{f\in \mbs{I}_{\tilde{F}}} \pdf_f^{(B)}\left(\mbs{w}_{n,f} | \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right).
            \end{equation}
            Where $\mbs{I}_{\tilde{F}}$ is the set of indices of the features, omitting missing phase space dimensions.

        % ssub: off_stream (end)

    % sub: astrometric_model (end)

@@

            At any given $\phi_1$, the isochrone is modeled in absolute magnitudes
            and parameterized by the normalized stellar mass, labeled $\gamma \in (0, 1]$, in the modeled mass range. For example, increasing $\gamma$ corresponds to movement up the main sequence towards the red giant branch. We denote the isochrone as $\mcal{I(\gamma)}$. The intrinsic dipsersion around the isochrone is $\Sigma_\mcal{I}(\gamma)$. Since the isochrone is in absolute magnitudes, but the data are in apparent magnitudes,
            it is necessary to shift the isochrone by a distance modulus to the predicted distance of the stream, labeled  $\mu(\phi_1)$, with intrinsic stream distance variation $\sigma_\mu(\gamma)$.

            For the $n$-th star, the photometric model is:
            \begin{multline} \label{eq:photometric_probability_point}
                \cmp{S,m}{P}(\mbs{m_i}, \gamma | \mbs{\theta}(\phi_1)) 
                \\ \sim \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1} \sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \\ \times \prior(\gamma, \nth{\mbs{m}}, \phi_1),
            \end{multline}

            where $\cmp{m}{\nth{\mbs{\Sigma}}}$ is the data covariance and $\prior(\gamma, \nth{\mbs{m}}, \phi_1)$ encodes both the present-day mass function of the stream, as well as any observational constraints (e.g. $g < 20$ mag).

            We are interested in the total probability (at some $\phi_1$) and so integrate over $\gamma$ to find:
            \begin{small}
            \begin{align} \label{eq:photometric_probability}
                &\cmp{S,m}{P}(\nth{\mbs{m}} | \mbs{\theta}(\phi_1)) = \\
                &\ \int\limits_{\gamma=0}^{1} \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1}\sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \nonumber
                \\
                &\qquad \times \prior(\gamma, \nth{\mbs{m}}, \phi_1) \rm{d}\gamma. \nonumber
            \end{align}
            \end{small}
            Then, for a given $\phi_1$ we can obtain a distribution over the distance modulus using 
            \begin{align}
                \mu_{\rm sample}\left(\phi_1\right) \sim \mathcal{N}\left(\mu(\phi_1), \sigma^2_\mu(\phi_1) \right) & \quad\unit{mag}.
            \end{align}
            For each sample, we can convert the distance-modulus distribution (and error) to the distance track using 
            \begin{align}
                \rm{dist}\left(\phi_1\right) &= 10^{\frac{1}{5}\mu_{\rm sample}(\phi_1)+1} & \quad\unit{pc}, \\
                \delta \rm{dist}(\phi_1) &\simeq \frac{\ln{10}}{5} \rm{dist}(\phi_1) \ \sigma_\mu(\phi_1) & \quad\unit{pc}.
            \end{align}

            Let $\cmp{S,m}{\mbs{\theta}}(\phi_1)$ represent the feed-forward neural network which outputs a vector of parameters characterizing the stream's photometric distribution at a given $\phi_1$. The network has 2 outputs -- $\mu, \Sigma_\mu$ -- as we only consider the distance track of the isochrone, holding fixed stellar population parameters internal to modeling the isochrone, such as its age and metallicity.
            In principle we might include parameters of the isochrone in the model, now $\mcal{I}(\mbs{\theta}, \gamma)$, where $\theta = {a, Z, ...}$ and any other generating parameters of the isochrone.
            % FIXME! DO this Likewise, the dust map may be treated as a prior and it's errors incorporated into the data errors.
            However, stellar streams are generally poorly fit by theoretic isochrones, at least ones with realistic ages and metallicities.
            Consequently including the age and metallicity as model parameters
            does not meaningfully contribute to constraining results on the stream's stellar population. Moreover, age and metallicity are partially degenerate with the distance modulus. An alternate approach is to use data-driven isochrones instead of theoretic ones. In some forms of this approach the age and metallicity are not meaningful quantities. In others, e.g. with a suite of models or a trained emulator, the data-driven isochrones might have realistic ages and metallicities and thus meaningfully contribute to the model.
            % Is this true? re-read Uniform Modeling of 13 streams
            We defer this to a future work.
